Self-Supervised Learning of Generalized Policies in multi-agent domains

### 2021-08-30 doc/paper/paper.tex
doc update

\usepackage[margin=1in]{geometry}

% our packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{bm}
\usepackage{booktabs}
%\usepackage[inline]{enumitem}
%\usepackage{mathtools}
%\usepackage{multirow}
%\usepackage[mode=buildnew,subpreambles=true]{standalone}
\usepackage{subcaption}
%\usepackage{todonotes}
\usepackage[table]{xcolor}  % TODO Comment out for final submission
%\usepackage{pgfplots}
%\usepackage{tikz}
\usepackage{natbib}
\usepackage{fancyvrb}
\usepackage{color}

@@

\title{Self-Supervised Learning of Generalized Policies in adversarial domains}
%Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
%\author{GFM}

@@

Generalized planning is concerned with the computation of general policies that solve multiple instances of a planning domain all at once. It has been recently shown how these policies can be computed and we adress the problem of computing general policies for multi-agent settings. These policies can be computed in two steps: first, a suitable abstraction in the form of a qualitative numerical planning problem (QNP) is learned from sample plans, then the general policies are obtained from the learned QNP using a planner. We introduce an alternative approach for computing more expressive general policies in multi-agent settings which does not require sample plans or a QNP planner. The new formulation is very simple and can be cast in terms that are more standard in machine learning: a large but finite pool of features (functions) is defined from the predicates in the planning examples using a general grammar, and a small subset of features is sought for separating “good” from “bad” state transitions, and goals from non-goals. The problems of finding such a “separating surface” while labeling the transitions as “good” or “bad” are jointly addressed as a single combinatorial optimization problem expressed as a Weighted Max-SAT problem. The advantage of looking for the simplest policy in the given feature space that solves the given examples, possibly non-optimally, is that many domains have no general policies that are optimal. We assume that the state space is discrete and also that we know what actions the opponents can take in their turn. The approach yields general policies for a number of benchmark adversarial domains. We use a uniform grammar for all the grid domains capable of capturing the domain representations. We introduce the possibility to use actions to reduce the number of features used to distinguish one transition from another, in case the action space does not change from one instance of the domain to another. In case the action space changes from one instance to another we simply ignore the actions used. These policies can be interpreted qualitatively, explained, and it can be shown mathematically that they will solve all possible instances within a class of problems. This is an advantage when making manual policy changes.

@@

Generalized planning is concerned with the computation of general policies or plans that solve multiple instances of a given planning domain all at once~\cite{srivastava2008network, ramirez2009plan, belle2016foundations, aguas2016generalized}. If the instances share the same strategy, it doesn't matter how different they are. For instance, in an environment where an agent has to pick up a key and then approach a door and it will open, the strategy is to approach the key, pick it up, and then approach the door in order to open it. In this example it does not matter how large the environment is and how many actions the agent has to take, since the strategy does not change.

The key question in generalized planning is how to represent and compute such general plans from the domain representation. This was solved by \cite{frances2021learning}, for single agent settings and deterministic domains but not for multi-agent settings, where you cannot control what the other agents will do during their turn. However, we deal with adversarial domains where the agent can take actions that no matter what happens in the environment the agent can win. For instance, if we are in a domain where an agent must chase an adversary, and the adversary can run slower than the agent, the strategy will be to decrease the distance between them until the agent can catch him. We know this is a winning strategy where it does not matter what the adversary does, the agent will win

In one of the most general formulations, general policies are obtained from an abstract planning model expressed as a qualitative numerical planning problem or QNP~\cite{srivastava2008network}. A QNP is a standard STRIPS planning model extended with non-negative numerical variables that can be decreased or increased “qualitatively”; i.e., by uncertain positive amounts, short of making the variables negative. Unlike standard planning with numerical variables~\cite{helmert2002decidability}, QNP planning is decidable, and QNPs can be compiled in polynomial time into fully observable non-deterministic (FOND) problems~\cite{bonet2020high}.

The main advantage of the formulation of generalized planning based on QNPs is that it applies to standard relational domains where the pool of (ground) actions change from instance to instance. However, in our apporach there is also the possibility to take advantage of actions if they do not change from one instance to another. On the other hand, while the planning domain is assumed to be given, the QNP abstraction is not, and hence it has to be written by hand or learned. This is the approach of~\cite{DBLP:conf/aaai/BonetFG19} where generalized plans are obtained by learning the QNP abstraction from the domain representation and sample plans, and then solving the abstraction with a QNP planner.

We take inspiration from the work mentioned above, but introduce an alternative approach for computing general policies in multi-agent setting that is simpler. The learning problem is cast as a self-supervised classification problem where (1) a pool of features is automatically generated from a uniform grammar applied to the domain predicates, and (2) a small subset of features is sought for separating “good” from “bad” state transitions, and goals from non-goals. The problems of finding the “separating surface” while labeling the transitions as “good” or “bad” are addressed jointly as a single combinatorial optimization task solved with a Weighted Max-SAT solver. The approach yields general policies for a number of benchmark domains. One of the advantages of having explainable policies is that it gives the engineer the possibility to modify them easily if is necessary. For instance, if there are known exceptions within the domain.

@@

\section{Contriution: Generalized Planning in multi-agent settings}

A key question in generalized planning is how to represent general plans or policies when the different instances to be solved have different sets of objects and ground actions. We take inspiration of the work of~\cite{frances2021learning} and we extend it to multi-agent domains. One solution is to work with general features (functions) that have well defined values over any state of any possible domain instance, and think of general policies $\pi$ as mappings from feature valuations into abstract actions that denote changes in the feature values~\cite{bandres2018planning}. When a transition is taken by the agent the environment or other agent changes state, so when we record the AND/OR graph of states we capture how the environment or the other agent can change each state. In this work, we use abstract actions as~\cite{frances2021learning}, in addition to the possibility of taking advantage of using an action space fixed for all instances in the domain.  The approach is SAT-based, so first we will define the variables and encoding, and then how we learn general policies.

@@

Let us have a pool of features $f$. Where $f$ is a function that has as input a state and as output a numerical value. The initial feature pool is produced using the logical description language. In addition, a parameter is used to determine the maximum complexity of the features. Once the initial pool of features is produced, a pruning is performed based on the denotation of each feature with the following philosophy: if two features, $f$ and $f'$ have the same denotation in all states, it is likely that they are capturing the same information of the domain, therefore we keep the simpler of the two features. With these features you can basically create abstract states, or representations, the challenge will be to find which ones are useful to represent the strategy that solves the problem. After feature pruning, the set of features $f$ that we are left with is $F$, and it is the one that will have the SAT solver as input. Based on the encoding the SAT solver will select the features of $F$ needed to represent the policy, trying to minimize the complexity.

@@

A state is called \emph{reachable} if there is a path to it from $s_0$, and it is called \emph{solvable} if there is a path from it to a goal state. A state is \emph{alive} if it is solvable, reachable and not a goal state~\cite{frances-et-al-ijcai2019}. We use $T$ to denote the set of all transitions $(s, s')$ in the training sample such that $s$ is alive.

@@

\subsubsection{Soft Constraints}
\smallpar{C6}
A constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to its complexity $\mathcal{K}(f)$.

@@

How we arrive at these constrains has to do with the observation of adversarial domains and winning strategies. We observed that the dilemma of adversarial games meant that we were confronted with the fact that decisions were not made by a single agent, and we had to take into account that there were also other agents trying to "win". Moreover, this meant that in many domains we were faced with scalability problems.

@@

We observed that in many domains the set of actions is fixed for all agents. In the paper~\cite{frances2021learning} there were $Good(s, s')$ variables. But these were not able to take advantage of fixed action sets, so we decided to include the information of the actions in these domains to alleviate scalability problems, for that reason we have variables such as $Good(s, a, s')$.
 This allows us to reduce the number of C4 clauses.

It was assumend by~\cite{frances2021learning} assumed that it was possible to get from any state of the problem to the objective, without entering dead ends (states where the agent has lost). However, in many multi-agent domains there are states where the agent has not lost, but no longer has the possibility of winning in that state no matter what the agent does. That is why we included the $Bad(s)$ variables, so that there could be states without any actions that lead to a path where the agent always wins. However, in our approach, if from a certain state there is a strategy that can be represented with the logic language and that always wins over the rest of the agents, we guarantee that we will find it. However, allowing the use of $Bad(s)$ variables is optional, through~$\texttt{allow\_bad\_states}$, since there are domains where we know that there is a winning strategy from any state.

In C1 it can be understood that each state must have a possible transition that is good, if ~$\neg\texttt{allow\_bad\_states}$. This makes sense for many of the domains, as we want our overall strategy to be able to decide which transition to take in any given state. We then design counterin C2 to ensure that the strategy takes paths that do not fall into cycles by introducing the adversarial factor. To introduce the adversarial factor into the strategy we make sure that the number assigned to a state $s$ is greater than all the numbers assigned to the possible $s''$. The possible $s''$ are the result of the agents or the environment executing an action $a'$ on $s'$, where $s'$ is the state the agent reaches after executing $a$ on $s$. Note that adversaries will often have multiple $a'$ actions to execute on $s'$. We then design counterin C2 to ensure that the strategy takes paths that do not fall into cycles by introducing the adversarial factor. To introduce the adversarial factor into the strategy we make sure that the number assigned to a state $s$ is greater than all the numbers assigned to the possible $s''$. The possible $s''$ are the result of the agents or the environment executing an action $a'$ on $s'$, where $s'$ is the state the agent reaches after executing $a$ on $s$. Note that adversaries will often have multiple $a'$ actions to execute on $s'$, which will often result in a different set of $s"$..

@@

\subsection{Optimizations}

\paragraph{Non-distinguishability of transitions as an equivalence relation.}
Any fixed, given pool of features $F$ implicitly defines an equivalence relation where to transitions are
equivalent iff they cannot be distinguished \emph{by any feature in $F$}.
If two transitions cannot be distinguished by any feature, then clearly either the policy computed by the SAT solver
considers all of them as ``good'', or as ``bad''.
We'll exploit this by using one single SAT variable to denote whether \emph{all transitions in a given equivalence
class} are good or bad. When exploiting this notion of equivalence (which is implemented as an optional feature of
the CNF generator), then every mention below to SAT variable $Good(s, s')$ needs to be read as $Good(s_{\star}, s_{\star}')$,
where $(s_{\star}, s_{\star}')$ is the \emph{representative} transition of the equivalence class to which $(s, s')$ belongs.

\paragraph{``Bad'' transitions.}
We use \badtx{} to denote the set of transitions that have been determined at preprocessing as necessarily
\emph{not} good.
At the moment, this set contains all transitions that go from an alive to an unsolvable state and, if using
equivalence relations, all those other transitions in whose equivalence class there is some other ``bad'' transition.

\section{Grammar}

The sorts used are: $cell$, $row$, $column$ and . The constants for the sort $cell$ have the form $cell.r\_c$, this constant refers to the cell that is at the same time in the row $r$ and in the column $c$. The constants for sort $row$ are $row.r$, and this constant refers to the row $r$. The constants for sort $column$ are $column.c$, and this constant refers to column $c$. Each sort can be linked to an object, using a one-arity predicate, depending on the objects of the domain, given by a predicate of arity one with any variable belonging to the sorts $cell$, $row$ or $column$. An example for the sort $cell$ would be $cell\_has\_v(cell)$, where the variable of type $cell$ can take the value of $cell.r\_c$, and this predicate would indicate that the cell $cell.r\_c$ has an object with value $v$, which is the name of the object currently in $cell.r\_c$. To map the relative position between constants of the same sort, we use two-arity predicates, which have as variables these constants belonging to the same sort. For example, $adjacent\_cell(cell, cell)$ takes two constants of the sort $cell$, and means that the two cells are adjacent. The same philosophy would apply to the predicates $adjacent\_col(column, column)$ or $adjacent\_row(row, row)$.

@@

@@

\subsubsection{Training instances:}
\begin{Verbatim}[fontsize=\footnotesize]
  # m  m  .  m #
  # m  m  .  m #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  A  . #
\end{Verbatim}

@@

    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ..m.mm # ..m.mm # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ..m.m. # ..m... # ..m.mm # ..m..m # ..m..m # ...... # ...... # ...... # ...... # ...... #
    # ..m... # ..m... # ..m... # ..m... # ..m... # ..m..m # ..m... # ..m... # ...m.. # ...m.. #
    # ...... # ...... # ..m... # ..m... # ..m... # ..m... # ..m... # ..m... # ...m.. # ...m.. #
    # ...... # ...... # ...... # ...... # ...... # ..m... # ..m... # ..m... # ...m.. # ...m.. #
    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
... # ....A. # ....A. # ....A. # ....A. # .....A # .....A # .....A # ....A. # ....A. # ...A.. # ...

@@

@@

\subsection{Dealing with real robots}

First it is necessary to have a simulator of the robot's environment discretised. The policy would be learned with the algorithm described in the paper without involving the robot. This policy is learned from a single instance in the simulator, and can then be run on different and larger instances of the domain. Once the policy is learned with the simulator and the algorithm, it is only necessary to translate the state where the robot is into a state that can be executed by the simulator. Then a transition is selected as good using the already learned policy, and the policy returns the action executed in the simulator to take that transition. The robot controller must be able to handle executing these actions in the real world. This process is repeated until the goal is reached.

So, to be able to execute the policies on real robots, it is only necessary to have a simulator and an algorithm to translate the state of the robot into a state executable by the simulator.

For example, with a turtle robot that has to pick up a certain number of packages and take them to a certain place. We only need a grid simulator, where the objects "package", "agent" and "destination" exist. The simulator must be able to execute the actions to move the robot, pick and drop. The strategy that we would learn with the simulator and our algorithm would be the same as the one learned in the "Delivery" domain, in the results section. The robot controller should be able to cope with executing the actions output by the algorithm, and ideally end up in the state that the simulator expects.

Of course, you could give the initial state translated to the simulator and have the learned policy execute a sequence of actions up to the goal. But this method makes us more vulnerable to the noise that the real world produces.

Note that for the algorithm to work it must be possible to win from the initial state of the problem no matter what the opponent does. So in domains where the agent has two actions per turn, which are "Space Invader" and "Delivery". If both players have one action per turn the agent's strategy does not change. Therefore in some domains the learned policy can deal with a noisy adversary even though it has not been learned by playing against him.

>>>>

\usepackage[margin=1in]{geometry}
\usepackage{csquotes}

% our packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{bm}
\usepackage{booktabs}
%\usepackage[inline]{enumitem}
%\usepackage{mathtools}
%\usepackage{multirow}
%\usepackage[mode=buildnew,subpreambles=true]{standalone}
\usepackage{subcaption}
%\usepackage{todonotes}
\usepackage[table]{xcolor}  % TODO Comment out for final submission
%\usepackage{pgfplots}
%\usepackage{tikz}
\usepackage{natbib}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{hyperref}

@@

\title{Self-Supervised Learning of Generalized Policies in multi-agent domains}
%Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
%\author{GFM}

@@

Generalized planning is concerned with the computation of general policies that solve multiple instances of a planning domain all at once. Recently, it has been recently shown how these policies can be computed for single-agent settings. These policies can be computed in two steps: first, a suitable abstraction in the form of a qualitative numerical planning problem (QNP) is learned from sample plans, then the general policies are obtained from the learned QNP using a planner. We introduce an alternative approach for computing more expressive general policies in multi-agent settings which does not require sample plans nor a QNP planner. The new formulation is very simple and can be cast in terms that are more standard in machine learning: a large but finite pool of features (functions) is defined from the predicates in the planning examples using a general grammar, and a small subset of features is sought for separating “good” from “bad” state transitions, as well as goals from non-goals. The problem of finding such a ``separating surface'' while labeling the transitions as ``good'' or ``bad'' are jointly addressed as a single combinatorial optimization problem expressed as a Weighted Max-SAT problem. As many domains have no general policies that are optimal for all instance, we argue that looking for a simple policy, which might have some performance reduction  is beneficial . We assume that the state space is discrete and also that we know what actions the agents can take in their turn. The approach yields general policies for a number of benchmark multi-agent domains. We use a uniform grammar for all the grid domains capable of capturing the domain representations. We introduce the possibility to use actions to reduce the number of features used to distinguish one transition from another, in case the action space does not change from one instance of the domain to another. In case the action space changes from one instance to another we simply ignore the actions used. We make use of the minimax algorithm to extend the meaning of an unsolvable state, ensuring that the agent wins from all states where the policy is defined, regardless of what the other agents do. These policies can be interpreted qualitatively, explained, and it can be shown mathematically that they will solve all possible instances within a class of problems. This is an advantage when making manual policy changes.

@@

Generalized planning is concerned with the computation of general policies or plans that solve multiple instances of a given planning domain all at once~\cite{srivastava2008network, ramirez2009plan, belle2016foundations, aguas2016generalized}. If the instances share the same strategy, it doesn't matter how different they are. For instance, in an environment where an agent has to pick up a key and then go to the door and it will open, the strategy is to approach the key, pick it up, and then approach the door in order to open it. In this example it does not matter how large the environment is and how many actions the agent has to take, since the strategy does not change.

The key question in generalized planning is how to represent and compute such general plans from the domain representation. This was solved by~\cite{frances2021learning}, for single agent settings and deterministic domains, where one agent cannot control what the other agents will do during their turn. We take inspiration of~\cite{frances2021learning} to create an approach that is capable of deal with multi-agent domains where the agent can take actions that no matter what happens in the environment the agent can win. For instance, if we are in a domain where an agent must chase an adversary, and the adversary can run slower than the agent, the strategy will be to decrease the distance between them until the agent can catch him. We know this is a winning strategy where it does not matter what the adversary does, the agent will win

In one of the most general formulations, general policies are obtained from an abstract planning model expressed as a qualitative numerical planning problem or QNP~\cite{srivastava2008network}. A QNP is a standard STRIPS planning model~\cite{fikes1971strips, bylander1994computational} extended with non-negative numerical variables that can be decreased or increased “qualitatively”; i.e., by uncertain positive amounts, short of making the variables negative. Unlike standard planning with numerical variables~\cite{helmert2002decidability}, QNP planning is decidable, and QNPs can be compiled in polynomial time into fully observable non-deterministic (FOND) problems~\cite{bonet2020high}.

The main advantage of the formulation of generalized planning based on QNPs is that it applies to standard relational domains where the pool of (ground) actions change from instance to instance. For instance, given a domain that is a grid where the actions are clicking a cell given some coordinates, the number of actions would be equal to the number of cells in the grid, then in this case the number of possible actions is different depending on the size and shape of the grid. On the other hand, while the planning domain is assumed to be given, the QNP abstraction is not, and hence it has to be written by hand or learned. This is the approach of~\cite{DBLP:conf/aaai/BonetFG19} where generalized plans are obtained by learning the QNP abstraction from the domain representation and sample plans, and then solving the abstraction with a QNP planner.

In our approach there is also the possibility to take advantage of actions if they do not change from one instance to another. For instance if the domain is a grid where the agent has to pick an object, the actions will be "move in some direction" and "pick", then the action space do not change with the size of the grid or the number of objects to be collected. The way we take advantage of domains with a fixed action space is by distinguishing with the features only the transitions that are the product of the same action, the mathematical expression of this effect is seen in the encoding section. To put it intuitively, let $tx=(s, a, s')$ be a transition from state $s$ to state $s'$ by executing the action $a$ and a transition $tx'=(t, b, t')$. If $a=b$ in a fixed action space, and $tx$ is a \emph{good} transition and $tx'$ is a \emph{bad} transition, we want to capture what change has occurred in the features. On the other hand, if $a$ is different from $b$ in a fixed action space we do not need to differentiate this transition.

We take inspiration from~\cite{frances2021learning}, but introduce an alternative approach for computing general policies in multi-agent settings that is simpler. The learning problem is cast as a self-supervised classification problem where (1) a pool of features is automatically generated from a uniform grammar applied to the domain without any domain knowledge, and (2) a small subset of features is sought for separating “good” from “bad” state transitions, and goals from non-goals. The problems of finding the final features while labeling the transitions as “good” or “bad” are addressed jointly as a single combinatorial optimization task solved with a Weighted Max-SAT solver. The approach yields general policies for a number of benchmark domains. One of the advantages of having explainable policies is that it gives the engineer the possibility to modify them easily if is necessary. For instance, if there are known exceptions within the domain.

The domain knowledge entered will come from the grammar used, which depending on the domains can be uniform. For instance, in this work we use domains based on grid worlds, which allow us to have a uniform grammar for all of them, having predicates for the value of the cells, and the relative position of some cells to others. Once the initial feature pool is automatically generated from a grammar, we do a pre-pruning of these features based on their denotation, without introducing domain knowledge. Of course, this pre-pruning is not sound, however we have observed that features with the same denotation often contain the same domain information. After pre-pruning the feature pool, a small set of features is selected automatically based on the encoding for representing the policy. The encogin guarantee that the number of features selected for representing the policy will be as small and with as little complexity as possible.

The approach is learning a model. The general policies obtained are made up of one or more rules, these rules have the form of "if a condition is met in certain features it is good to take actions that change the features in this way". This is completely different from the most common approaches, as it is not necessary to have any explicit action for the policy to be defined. However, if there is a fixed set of actions the approach is able to take advantage of this fact. Moreover, this way of using the features to represent the policy allows us to generalise to larger and more complex instances, since we are capturing the strategy that solves the problem.

In~\cite{frances2021learning}, an \emph{unsolvabel} state is defined as a state in which the agent has lost. But in multi-agent games the agent may find himself in a state where he has not lost and yet cannot win in any way if the adversary or adversaries act correctly. Therefore, we also make use of the minimax algorithm to extend the meaning of an unsolvable state, ensuring that the agent wins from all states where the policy is defined, regardless of what the other agents do. The extension of unsolvable: An state is unsolvabel if some adversary is able to win regardless of what the agent does.

Inspiration has been taken from the article by~\cite{silver2020few} to choose the domains. In~\cite{silver2020few} interpretable policies were learned by expanding with expert policies and then inferring the expert policy with the target language. In other words, imitation learning was used and the policies were learned in a supervised way. In this work we learn policies for the games of~\cite{silver2020few}, such as "two pile nim", "checkmate tactic" or "chase", but in an unsupervised way.

@@

\section{Contribution: Generalized Planning in multi-agent settings}

A key question in generalized planning is how to represent general plans or policies when the different instances to be solved have different sets of objects and ground actions. We take inspiration of the work of~\cite{frances2021learning} and we extend it to multi-agent domains. One solution is to work with general features (functions) that have well defined values over any state of any possible domain instance, and think of general policies $\pi$ as mappings from feature valuations into abstract actions that denote changes in the feature values~\cite{bandres2018planning}. When a transition is taken by the agent the environment or other agent changes state, so when we record the AND/OR graph of states, capturing how the environment or the other agent can change each state. In this work, we use abstract actions as~\cite{frances2021learning}, in addition to the possibility of taking advantage of using an action space fixed for all instances in the domain.  The approach is SAT-based, so first we will define the variables and encoding, and then how we learn general policies.

@@

Let us have a pool of features $f$. Where $f$ is a function that has as input a state and as output a numerical value. The initial feature pool is produced using the logical description language, i.e., genereted from a grammar. In addition, a parameter is used to determine the maximum complexity of the features. Once the initial pool of features is produced, a pruning is performed based on the denotation of each feature with the following philosophy: if two features, $f$ and $f'$ have the same denotation in all states, it is likely that they are capturing the same information of the domain, therefore we keep the simpler of the two features. With these features we can basically create abstract states, or representations, the challenge will be to find which ones are useful to represent the strategy that solves the problem. After the feature pruning, the set of features $f$ that are left represent $F$, and it is the one that will have the SAT solver as input. Based on the encoding the SAT solver will select the features of $F$ needed to represent the policy, trying to minimize the complexity and the number of features used.

@@

A state is called \emph{reachable} if there is a path to it from $s_0$, and it is called \emph{solvable} if there is a path from it to a goal state. A state is \emph{alive} if it is solvable, reachable and not a goal state~\cite{frances-et-al-ijcai2019}. A state is \emph{unsolvable} if the agent has lost. We use $T$ to denote the set of all transitions $(s, a, s')$ in the training sample such that $s$ is alive.

\subsubsection{Extension of unsolvaility}

In some domains the agent can win from any state of the problem except for some specific states where the agent has lost. In other more complex domains, there are states where the agent has not lost, however, the success of the agent depends on whether the opponent plays well or not. In other words, if the opponent plays correctly the agent has no way to win.

With our approach we want policies that are always able to win, no matter what state the agent is in. Therefore, we consider states where the agent can lose no matter what he does as \emph{unsolvable}. To classify unsolvable states we use the minimax algorithm.

@@

\subsubsection{Soft Constraints}

\smallpar{C6}
A constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to its complexity $\mathcal{K}(f)$.

@@

How we arrive at these constrains has to do with the observation of multi-agent domains and winning strategies. We observed that the dilemma of adversarial games meant that we were confronted with the fact that decisions were not made by a single agent, and we had to take into account that there were also other agents trying to "win". Moreover, this meant that in many domains we were faced with scalability problems.

@@

We observed that in many domains the set of actions is fixed for all agents. In the paper~\cite{frances2021learning} there were $Good(s, a, s')$ variables. But these were not able to take advantage of fixed action sets, so we decided to include the information of the actions in these domains to alleviate scalability problems, for that reason we have variables such as $Good(s, a, s')$.
 This allows us to reduce the number of C4 clauses.

It was assumend by~\cite{frances2021learning} assumed that it was possible to get from any state of the problem to the objective, without entering dead ends (states where the agent can lose, independently of what he does). However, in many multi-agent domains there are states where the agent has not lost, but no longer has the possibility of winning in that state no matter what the agent does. That is why we included the $Bad(s)$ variables, so that there could be states without any actions that lead to a path where the agent always wins. However, in our approach, if from a certain state there is a strategy that can be represented with the logic language and that always wins over the rest of the agents, we guarantee that we will find it. However, allowing the use of $Bad(s)$ variables is optional, through~$\texttt{allow\_bad\_states}$, since there are domains where we know that there is a winning strategy from any state.

In C1 it can be understood that each state must have a possible transition that is good, if ~$\neg\texttt{allow\_bad\_states}$. This makes sense for many of the domains, as we want our overall strategy to be able to decide which transition to take in any given state. We then design counterin C2 to ensure that the strategy takes paths that do not fall into cycles by introducing the adversarial factor. To introduce the adversarial factor into the strategy we make sure that the potential $V$ of the state $s$ is greater than all the potentials $V$ of all the possible $s''$. The possible $s''$ are the result of the agents or the environment executing an action $a'$ on $s'$, where $s'$ is the state the agent reaches after executing $a$ on $s$. Note that adversaries will often have multiple $a'$ actions to execute on $s'$, which will often result in a different set of $s''$.

@@

%\subsection{Optimizations}

%\paragraph{Non-distinguishability of transitions as an equivalence relation.}
%Any fixed, given pool of features $F$ implicitly defines an equivalence relation where to transitions are
%equivalent iff they cannot be distinguished \emph{by any feature in $F$}.
%If two transitions cannot be distinguished by any feature, then clearly either the policy computed by the SAT solver
%considers all of them as ``good'', or as ``bad''.
%We'll exploit this by using one single SAT variable to denote whether \emph{all transitions in a given equivalence
%class} are good or bad. When exploiting this notion of equivalence (which is implemented as an optional feature of
%the CNF generator), then every mention below to SAT variable $Good(s, a, s')$ needs to be read as $Good(s_{\star}, s_{\star}')$,
%where $(s_{\star}, a, s_{\star}')$ is the \emph{representative} transition of the equivalence class to which $(s, a, s')$ belongs.

%\paragraph{``Bad'' transitions.}
%We use \badtx{} to denote the set of transitions that have been determined at preprocessing as necessarily
%\emph{not} good.
%At the moment, this set contains all transitions that go from an alive to an unsolvable state and, if using
%equivalence relations, all those other transitions in whose equivalence class there is some other ``bad'' transition.

\section{Motivating Example}

We will now consider a grid-world in which the agent has to reach a goal. The grid-world is made of cell, organized in rows and columns. The sorts used are: $cell$, $row$, $column$ and . The constants for the sort $cell$ have the form $cell.r\_c$, this constant refers to the cell that is at the same time in the row $r$ and in the column $c$. The constants for sort $row$ are $row.r$, and this constant refers to the row $r$. The constants for sort $column$ are $column.c$, and this constant refers to column $c$. Each sort can be linked to an object, using a one-arity predicate, depending on the objects of the domain, given by a predicate of arity one with any variable belonging to the sorts $cell$, $row$ or $column$. An example for the sort $cell$ would be $cell\_has\_v(cell)$, where the variable of type $cell$ can take the value of $cell.r\_c$, and this predicate would indicate that the cell $cell.r\_c$ has an object with value $v$, which is the name of the object currently in $cell.r\_c$. To map the relative position between constants of the same sort, we use two-arity predicates, which have as variables these constants belonging to the same sort. For example, $adjacent\_cell(cell, cell)$ takes two constants of the sort $cell$, and means that the two cells are adjacent. The same philosophy would apply to the predicates $adjacent\_col(column, column)$ or $adjacent\_row(row, row)$. $cell\_has\_v()$ and $adjacent\_row/col()$ represent domain knowledge that needs to be encoded at design time.

\section{Using the algorithm}

In this section we describe the end-to-end process of the steps necessary to use the algorithm~\footnote{The code of our project: ~\url{https://github.com/iOrb/gpl}}. To run it we must first have a simulator of the domain in which we can execute an action and see the next state. We also have to take into account that we want to expand the whole state space, so we have to know to which states the adveraries can go.

\begin{enumerate}
  \item First we must decide which predicates we are going to use, e.g., $cell\_has\_v()$ . Then we will have to define these predicates programmatically, we have done it with Tarski~\footnote{Tarski is a framework for the specification, modeling and manipulation of AI planning problems. Code:~\url{https://github.com/aig-upf/tarski}}. This defines our language.

  \item Then we need to be able to transform any state of the problem into a set of atoms taking into account our predicates.

  \item Finally, we define the experiment, introducing the configuration parameters of the simulation environment if any. We can also modify the parameters of our approach, for example the maximum complexity of the features, which are generated incrementally based on the defined language. And we are ready to execute it, obtain the policy and test it in diferent instances.
\end{enumerate}

@@

Examples of games in which the learned policies are executed against other agents can be found in the appendix of the article.

@@

\subsubsection{Training instances}
\begin{Verbatim}[fontsize=\footnotesize]
  # m  m  .  m #
  # m  m  .  m #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  A  . #
\end{Verbatim}

@@

@@

%\subsection{Dealing with real robots}

%First it is necessary to have a simulator of the robot's environment discretised. The policy would be learned with the algorithm described in the paper without involving the robot. This policy is learned from a single instance in the simulator, and can then be run on different and larger instances of the domain. Once the policy is learned with the simulator and the algorithm, it is only necessary to translate the state where the robot is into a state that can be executed by the simulator. Then a transition is selected as good using the already learned policy, and the policy returns the action executed in the simulator to take that transition. The robot controller must be able to handle executing these actions in the real world. This process is repeated until the goal is reached.

%So, to be able to execute the policies on real robots, it is only necessary to have a simulator and an algorithm to translate the state of the robot into a state executable by the simulator.

%For example, with a turtle robot that has to pick up a certain number of packages and take them to a certain place. We only need a grid sifmulator, where the objects "package", "agent" and "destination" exist. The simulator must be able to execute the actions to move the robot, pick and drop. The strategy that we would learn with the simulator and our algorithm would be the same as the one learned in the "Delivery" domain, in the results section. The robot controller should be able to cope with executing the actions output by the algorithm, and ideally end up in the state that the simulator expects.

%Of course, you could give the initial state translated to the simulator and have the learned policy execute a sequence of actions up to the goal. But this method makes us more vulnerable to the noise that the real world produces.

Notice that for the algorithm to work it must be possible to win from the initial state of the problem no matter what the opponent does. So in domains where the agent has two actions per turn, which are "Space Invader" and "Delivery". If both players have one action per turn the agent's strategy does not change. Therefore in some domains the learned policy can deal with a noisy adversary even though it has not been learned by playing against him.

It is possible to learn policies unsupervised in small instances and transfer them to larger instances in the same domain. We can conclude from this that it is correctly learning the problem representation and a strategy capable of winning on a number of instances. Depending on the domains it is possible to create a common language for all, making the approach less domain dependent, although as we see in this work the language is focused on problems that are in two-dimensional grids. A requirement for the algorithm to work correctly is that the strategy can be described using the grammar, then in domains where the strategy is very complex it will be necessary to increase the complexity of the language. In this paper we propose a method for dealing with multi-agent domains, which manages to find policies that beat the adversary from any initial state where this is possible. This approach is a step in the explainability of reinforcement-produced policies, and it is also a step towards having transferable policies that capture the representation of the problem.

@@

\section{Apendix}

Here we show traces of the example domains, in which the learned policy is executed on small instances, against other agents and on larger instances.

Recall that the objects that can be found in the layouts are equivalent to:
\begin{Verbatim}[fontsize=\footnotesize]
  A: Agent
  D: destination
  T: adversary
  K: white-king
  Q: white-queen
  R: white-rook
  k: black-king
  m: martian
  p: pet
  t: token
\end{Verbatim}

\subsubsection{Training instances}
\begin{Verbatim}[fontsize=\footnotesize]
  # m  m  .  m #
  # m  m  .  m #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  .  . #
  # .  .  A  . #
\end{Verbatim}

    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ..m.mm # ..m.mm # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ..m.m. # ..m... # ..m.mm # ..m..m # ..m..m # ...... # ...... # ...... # ...... # ...... #
    # ..m... # ..m... # ..m... # ..m... # ..m... # ..m..m # ..m... # ..m... # ...m.. # ...m.. #
    # ...... # ...... # ..m... # ..m... # ..m... # ..m... # ..m... # ..m... # ...m.. # ...m.. #
    # ...... # ...... # ...... # ...... # ...... # ..m... # ..m... # ..m... # ...m.. # ...m.. #
    # ...... # ...... # ...... # .ee..... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
    # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... # ...... #
... # ....A. # ....A. # ....A. # ....A. # .....A # .....A # .....A # ....A. # ....A. # ...A.. # ...

\subsection{Nim}

\subsection{Delivery}

\subsection{Shoot}

\subsection{Checkmate in one with Queen}

\subsection{Checkmate in one with Rook}
